A Classical Density-Functional Theory for Describing Water Interfaces

### 2013-09-02 papers/water-saft/paper.tex
address fixmes Tested everything on bingley.

\subsection{Dispersion}

\fixme{Explain the dispersion term here.}

@@

\fixme{Talk about Fig.~\ref{fig:single-rod-broken-HB}}

>>>>

\subsection{Dispersion}
The dispersion free energy includes the van der Waals attraction and
any orientation-independent interactions. Following \hughesetal, we
use a dispersion term based on the SAFT-VR
approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
parameters (taken from Clark~\emph{et al}\cite{clark2006developing}):
an interaction energy $\epsilondisp$ and a length scale $\lambdadisp
R$.

The SAFT-VR dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
\begin{align}
  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
\end{align}
where $a_1$ and $a_2$ are the first two terms in a high-temperature
perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
the mean-field dispersion interaction. The second term, $a_2$, describes the
effect of fluctuations resulting from compression of the fluid due
to the dispersion interaction itself, and is approximated
using the local compressibility approximation (LCA), which
assumes the energy fluctuation is simply related to the
compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.

The form of $a_1$ and $a_2$ for SAFT-VR is given in
reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
of the packing fraction.  In order to apply this form to an
\emph{inhomogeneous} density distribution, we construct an effective local
packing fraction for dispersion $\etadisp$, given by a Gaussian
convolution of the density:
\begin{align}
  \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
  \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
      \lscale R)^2}}\right)d\xx'.\label{eq:packing-fraction}
\end{align}
This effective packing fraction is used throughout the dispersion
functional, and represents a packing fraction averaged over the
effective range of the dispersive interaction.
Eq.~\ref{eq:packing-fraction} contains an additional empirical
parameter $\lscale$ introduced by \hughesetal, which modifies the
length scale over which the dispersion interaction is correlated.

@@

In addition to the density, we examine the number of hydrogen bonds
which are broken due to the presence of a hard rod.  We define this
quantity as
\begin{align}
  N_{\text{broken HB}} &= 2 \int (X(\rr) - X_{\text{bulk}})n_{\text{site}}(\rr) d\rr
\end{align}
where $X_{\text{bulk}} = 0.1?$ \fixme{get this number} is the fraction
of unbonded association sites in the bulk.  The factor of 2 is chosen
to account for the four association sites per molecule, and the fact
that each hydrogen bond is twice---once for each of the molecules
involved.  In Fig.~\ref{fig:single-rod-broken-HB} we show the number
of hydrogen bonds broken by a hard rod per nanometer length, as
predicted by the functional of \hughesetal\ (dashed line) and this
work (solid line), as a function of the radius of the hard rod.  In
each case in the limit of large rods, the number of broken bonds is
proportional to the surface area, which is what one would expect.  At
every radius, the functional of \hughesetal\ predicts approximately
four times as many broken hydrogen bonds as the improved functional.
